Measuring contrast processing in the visual system using SSVEP

### 2024-04-20 review.qmd
Added new figures. Removed some html support files

@@

```{python}
#| label: fig-centresurround
#| fig-cap: "The perceived contrast of a stimulus depends on its context. A high contrast surround reduces the apparent contrast of the central 'probe' region."
#| fig-cap-location: bottom

@@

@@

Recent work on understanding Parkinson's disease has also used *Drosophila* genetic models. @Afsari2014 found that mutant flies produced stronger SSVEP responses to flickering lights than control flies (see [@fig-plotclinical]d). The authors theorised that differences in early gain control during development might lead to visual deficits later in life. Although visual responses are a convenient assay of neural function, it is likely that same general process applies throughout the whole brain, including in the motor system where the core Parkinson's symptoms manifest. The SSVEP differences were reduced by a kinase inhibitor that targets the dopamine system, demonstrating how model organisms can be used to test new pharmacological treatments. SSVEP responses also provide a potential method to diagnose Parkinson's before any symptoms manifest, and to monitor the effect of treatments.

@@

>>>>

```{python setup}

# all package loading and subfunctions defined here
import numpy as np
import matplotlib.pyplot as plt
from matplotlib import colormaps
import scipy as scp

def doubleTransducerFunction(inputContrast, maskContrast, c50, RMax):
 # This version has two inputs, but the second input is purely suppressive
  expnt=2
  inputContrast=np.abs(inputContrast)
  maskContrast=np.abs(maskContrast)

```{python createcentresurround}

plt.tight_layout()
plot.savefig('Figures/centresurround.pdf')

```{python linearsystem}

inputFrequency_Hz=np.array([5]) # We simulate a system with on input at many different contrasts
contrastRange=np.logspace(0,1.5,20)/100 # Equally spaced log contrast between 1 and 10^1.5
duration_S=1 # All variables in this code have an _[units] where possible
digitisationRate_Hz=1000
freqCutoff_Hz=30
c50=.10
RMax=10

   # Now compute the FT on the signal after it has passed through a single non-linearity. Do this 
   # using the transducer functiona above.
  transducedModulation=simpleTransducerFunction(inputModulation, c50,RMax)
  FT_transducedModulation=np.fft.fft(transducedModulation)/ len(transducedModulation)
  transducedAmplitude = np.abs(FT_transducedModulation[inputFrequency_Hz*2])
  transducedSignal[contrastIndex]=transducedAmplitude[0]

frequencySupport = np.arange(1,freqCutoff_Hz*duration_S+1) # We only want to plot the frequencies in the FFT up to this point

axs[0].plot(support_S, inputModulation)
axs[0].set_title('Input sine wave', fontsize=10)

# Plot the absolute values of the FFT in the second axis as a bar plot
axs[1].bar(frequencySupport, np.abs(FT_Signal[frequencySupport]))
axs[1].set_xlabel('Frequency (Hz)')
axs[1].set_ylabel('Magnitude (a.u.)')
axs[1].set_title('Fourier spectrum')
axs[1].grid()

axs[2].plot(contrastRange,initialSignal)
axs[2].set_xlabel('Contrast')
axs[2].set_ylabel('Amplitude (a.u.)')
axs[2].set_title('Linear computation stage', fontsize=10)
axs[2].grid()

axs[2].grid()
plt.tight_layout()
plot.savefig('Figures/linearsystem.pdf')

```{python transducerout}

   # Now compute the FT on the signal after it has passed through a single non-linearity. Do this 
   # using the transducer functiona above.
  transducedModulation=simpleTransducerFunction(inputModulation, c50,RMax)
  FT_transducedModulation=np.fft.fft(transducedModulation)/ len(transducedModulation)
  transducedAmplitude = np.abs(FT_transducedModulation[inputFrequency_Hz*2])
  transducedSignal[contrastIndex]=transducedAmplitude[0]

frequencySupport = np.arange(1,freqCutoff_Hz*duration_S+1) # We only want to plot the frequencies in the FFT up to this point

axs[0].plot(support_S, transducedModulation)
axs[0].set_title('Transduced signal', fontsize=10)

# Plot the absolute values of the FFT in the second axis as a bar plot
axs[1].bar(frequencySupport, np.abs(FT_transducedModulation[frequencySupport]))
axs[1].set_xlabel('Frequency (Hz)')
axs[1].set_ylabel('Magnitude (a.u.)')
axs[1].set_title('Fourier spectrum')
axs[1].grid()

axs[2].plot(contrastRange,transducedSignal)
axs[2].set_xlabel('Contrast')
axs[2].set_ylabel('Amplitude (a.u.)')
axs[2].set_title('Transducer output', fontsize=10)
axs[2].grid()

axs[2].grid()
plt.tight_layout()
plot.savefig('Figures/transducer.pdf')

```{python twoinputsout}

# Define some constants
inputFrequency_Hz=np.array([5]) # We simulate a system with on input at many different contrasts
contrastRange=np.logspace(0,1.5,20)/100 # Equally spaced log contrast between 1 and 10^1.5
maskFrequency_Hz=np.array([7]) # We simulate a system with on input at many different contrasts
maskContrast=np.array(0.3) # contrast of mask signal

duration_S=1 # All variables in this code have an _[units] where possible
digitisationRate_Hz=1000
freqCutoff_Hz=30
c50=.10
RMax=10

initialSignal = np.zeros(len(contrastRange)) 
transducedSignal = np.zeros(len(contrastRange))
maskedSignal = np.zeros(len(contrastRange))
for contrastIndex,thisContrast in enumerate(contrastRange):
  inputModulation = np.sin(2 * np.pi * inputFrequency_Hz * support_S) * thisContrast
  maskModulation = np.sin(2 * np.pi * maskFrequency_Hz * support_S) * maskContrast

   # Now compute the FT on the signal after it has passed through a single non-linearity. Do this 
   # using the transducer functiona above.
  transducedModulation=doubleTransducerFunction(inputModulation, 0*maskModulation, c50,RMax)
  FT_transducedModulation=np.fft.fft(transducedModulation)/ len(transducedModulation)
  transducedAmplitude = np.abs(FT_transducedModulation[inputFrequency_Hz*2])
  transducedSignal[contrastIndex]=transducedAmplitude[0]

  transducedModulation=doubleTransducerFunction(inputModulation, maskModulation, c50,RMax)
  FT_transducedModulation=np.fft.fft(transducedModulation)/ len(transducedModulation)
  transducedAmplitude = np.abs(FT_transducedModulation[inputFrequency_Hz*2])
  maskedSignal[contrastIndex]=transducedAmplitude[0]

frequencySupport = np.arange(1,freqCutoff_Hz*duration_S+1) # We only want to plot the frequencies in the FFT up to this point

axs[0].plot(support_S, transducedModulation)
axs[0].set_title('Transduced signal', fontsize=10)

# Plot the absolute values of the FFT in the second axis as a bar plot
axs[1].bar(frequencySupport, np.abs(FT_transducedModulation[frequencySupport]))
axs[1].set_xlabel('Frequency (Hz)')
axs[1].set_ylabel('Magnitude (a.u.)')
axs[1].set_title('Fourier spectrum')
axs[1].grid()

axs[2].plot(contrastRange,transducedSignal)
axs[2].plot(contrastRange,maskedSignal)
axs[2].set_xlabel('Contrast')
axs[2].set_ylabel('Amplitude (a.u.)')
axs[2].set_title('Transducer output', fontsize=10)
axs[2].grid()

axs[2].grid()
plt.tight_layout()
plot.savefig('Figures/twoinputs.pdf')

@@

![The perceived contrast of a stimulus depends on its context. A high contrast surround reduces the apparent contrast of the central 'probe' region.](Figures/centresurround.pdf){#fig-centresurround}

@@

![Linear.](Figures/linearsystem.pdf){#fig-linear}

![Linear.](Figures/transducer.pdf){#fig-transducer}

![Linear.](Figures/twoinputs.pdf){#fig-twoinputs}

@@

Recent work on understanding Parkinson's disease has also used *Drosophila* genetic models. @Afsari2014 found that mutant flies produced stronger SSVEP responses to flickering lights than control flies (see [@fig-plotclinical]d). The authors theorised that differences in early gain control during development might lead to visual deficits later in life. Although visual responses are a convenient assay of neural function, it is likely that same general process applies throughout the whole brain, including in the motor system where the core Parkinson's symptoms (tremor, rigidity, slow movement) manifest. The SSVEP differences were reduced by a kinase inhibitor that targets the dopamine system, demonstrating how model organisms can be used to test new pharmacological treatments. SSVEP responses also provide a potential method to diagnose Parkinson's before any symptoms manifest, and to monitor the effect of treatments.

@@

I wonder if this is a good place to mention attention.
